A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-18 tex/codatadp.tex
more

To construct a completion of $\Val$, we must choose values for the free variables of $\Val$, $v_{\ec{x}}$ and $v_{\ec{w_3}}$.
The content of set $\mathcal{T}^\M_{v_{\ec{x}}}( \Val )$ is $\alpha$-equivalent to
\[\begin{array}{c}
\{
\mu x. \const{A}( \mu y. \const{B}( x )),
\mu x. \const{A}( \mu y. \const{B}( y )),
\mu x. \const{B}( x ),
\mu x. \const{C}( \mu y. \const{A}( x )),
\mu x. \const{A}( \mu y. \const{C}( x )),
\mu x. \const{A}( x )
\}
\end{array}\]
Say that $\M$ interprets variables in $\Ec$ based on the mapping $\Val$,
for instance, $x$ is interpretted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$ and so on.
Assigning a value for $v_{\ec{x}}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^\M_{v_{\ec{x}}}( \Val )$
may cause values in the range of $\Val$ to become $\alpha$-equivalent, 
which in turn may cause $\M$ to be inconsistent.
For instance, say we assign $\mu v_{\ec{x}}. \const{B}( v_{\ec{x}} )$ for $v_{\ec{x}}$.
After this substitution, $\Val \ec{y_2}$ is $\mu v_{\ec{y_2}}. \const{B}( \mu v_{\ec{x}}. \const{B}( v_{\ec{x}} ) )$, 
which has normal form $\mu v_{\ec{y_2}}. \const{B}( v_{\ec{y_2}} )$,
which is $\alpha$-equivalent to $\mu v_{\ec{x}}. \const{B}( v_{\ec{x}} )$.
However, this contradicts the disequality $x \not\teq y_2$ in $\Ec$.

@@

The results are summarized in Table~\ref{tab:bench}. The decision procedure
makes a difference across all theories, and an overall difference of
\BAD{4.7\%}. Moreover, every aspect of the procedure, including the more
expensive rules, make a contribution. \textbf{NOT TRUE}: What is not visible in the table, but
readily available when looking at the raw data \cite{our-eval-data}, is that
the stronger decision procedures subsume the weaker ones. In the context of
Sledgehammer, the power of interpreted (co)datatype is comparable with that of
arithmetic \cite{blanchette-et-al-2013-smt}.

@@

>>>>

To construct a completion of $\Val$, we must choose values for the free variables of $\Val$, $v_{\ec{x}}$ and $v_{\ec{w_3}}$.
The content of set $\mathcal{T}^\M_{v_{\ec{x}}}( \Val )$ is $\alpha$-equivalent to
\[%\begin{array}{c}
\{
\mu x. \const{A}( \mu y. \const{B}( x )),
\mu x. \const{A}( \mu y. \const{B}( y )),
\mu x. \const{B}( x ),
\mu x. \const{C}( \mu y. \const{A}( x )),
\mu x. \const{A}( \mu y. \const{C}( x )),
\mu x. \const{A}( x )
\}
%\end{array}
\]
Say that $\M$ interprets variables in $\Ec$ based on the mapping $\Val$,
for instance, $x$ is interpretted as $\Val \ec{x}$, $y_1$ as $\Val \ec{y_1}$ and so on.
Assigning a value for $v_{\ec{x}}$ that is $\alpha$-equivalent to a value in $\mathcal{T}^\M_{v_{\ec{x}}}( \Val )$
may cause values in the range of $\Val$ to become $\alpha$-equivalent, 
which in turn may cause $\M$ to be inconsistent.
For instance, say we assign $\mu v_{\ec{x}}. \const{B}( v_{\ec{x}} )$ for $v_{\ec{x}}$.
After this substitution, $\Val \ec{y_2}$ is $\mu v_{\ec{y_2}}. \const{B}( \mu v_{\ec{x}}. \const{B}( v_{\ec{x}} ) )$, 
which has normal form $\mu v_{\ec{y_2}}. \const{B}( v_{\ec{y_2}} )$,
which is $\alpha$-equivalent to $\mu v_{\ec{x}}. \const{B}( v_{\ec{x}} )$.
However, this contradicts the disequality $x \not\teq y_2$ in $\Ec$.

@@

The results are summarized in Table~\ref{tab:bench}. The decision procedure
makes a difference across all theories, and an overall difference of
\BAD{4.7\%}. Moreover, every aspect of the procedure, including the more
expensive rules, make a contribution. \textbf{NOT TRUE}: What is not visible in the table, but
readily available when looking at the raw data \cite{our-eval-data}, is that
the stronger decision procedures subsume the weaker ones. In the context of
Sledgehammer, the power of interpreted (co)datatypes is roughly
comparable to that of arithmetic \cite{blanchette-et-al-2013-smt}.

@@

\begin{verbatim}
primcorec FIXME_x :: "nat tree" where
  "FIXME_x = Node 0 num FIXME_x"

lemma FIXME_x_unique: "x = Node 0 num x \<Longrightarrow> x = FIXME_x"
proof (coinduction arbitrary: x rule: tree.coinduct_strong)
  case (Eq_tree x)
  show ?case
    by (subst (1 2 3 4) Eq_tree) (simp add: eqTrueI[OF Eq_tree])
qed
\end{verbatim}
